% Copyright (C) Will Robertson 2004. 
%
% Permission is given to freely distribute and modify 
% this document provided attribution is given.

\documentclass[12pt,article]{memoir}

\usepackage{type1cm} % this is necessary to scale the dropped capital to the correct arbitrary size
\usepackage{amssymb} % for mathbb

\usepackage{lettrine} % the package that does all the work!

\setlength{\parindent}{0pt}
\setlength{\parskip}{\baselineskip}

\setlength{\DefaultNindent}{0pt}
\setcounter{DefaultLines}{2} % try 3 or 4!

\renewcommand{\LettrineFontHook}{\fontfamily{ppl}}

\begin{document}

\section*{Drop-caps with \textsf{lettrine}}

Some basic drop cap examples. \textit{Nota bene} You should globally set the font used for your drop caps with the \verb|\LettrineFontHook|. \textit{Id est} use a command like:\\
\verb|\renewcommand{\LettrineFontHook}{\fontfamily{ppl}}|\\
which changes the font used for the drop cap to (in this case) Palatino.

\lettrine{W}{elcome} to an example drop-capped paragraph. I want it really long so it extends over several lines, because there's really not much point in writing an example paragraph for a dropped capital if there is nowhere for it to protrude\dots.

\lettrine{\textsf{S}}{}% using a custom bit of font changing to get get smaller sans serif
{\sffamily\scriptsize ANS} serif doesn't look too bad. It's nice with the constant width `S'. The \textsf{lettrine} documentation contains a (French) demo of the advanced features of this package. With tweaking per character, you can get some fantastic results.

\lettrine{$\mathbb{P}$}{lease} have a look at the \texttt{fancy-math-fonts.tex} example for some more ideas on using alternate fonts here. I'd recommend Euler Caligraphic (with the \textsf{eucal} package) and Euler Fractur (with the \textsf{amssymb} or \textsf{amsfonts} packages).

\lettrine[loversize=0.2,findent=3pt,nindent=-2pt]{\fontfamily{pzc}\selectfont {R}}{emember}
% Zapf chancery needs some massaging to fit nicely---there are lots of facilities for this
 not to over-do any font changing in your document. I'm not even sure how suitable drop caps are in most circumstances. But don't they look nice?

\end{document}